---
title: "COVID-19 visualization"
output: html_document
---

If this is your first time in RMarkdown, welcome!
Some quick primers:
- To run a *line* of code, place your cursor there and `Cmd+Enter` (or `Ctrl+Enter` in Windows/Linux).
- To run a *chunk* of code, place your cursor there and `Cmd+Shift+Enter` (or `Ctrl+Shift+Enter` in Windows/Linux).
- To ask what a function does, try putting a `?` before the function name.
For example:
```{r}
?mean
```

I have put some lengthy script in `utils.R` so we only have to "source" it here.
This file contains `ft_theme()` (see Exercises) and 
the `selected_cou_FT` vector containing the country names we're interested in.

## Setup
```{r}
library(tidyverse)  
library(zoo) # calculate moving average
library(gghighlight)
library(coronavirus)

source('utils.R', echo=TRUE)
```

## Read in data

```{r}
# update_dataset()
data("coronavirus")
coronavirus %>% 
  head()
```

##  PART 1: Exploring

Choose a few countries to plot (or go with preselected ones):

```{r}
# selected_countries <- sample(unique(coronavirus$country), 5)
selected_countries <- selected_cou_FT
```

*Tip*: Piping into {.} allows you to comment out any line of the pipes
(including the last one).

```{r}
mov_avg_covid <- coronavirus %>%
  filter(country %in% selected_countries,
         type == "death") %>%
  group_by(date, country) %>%
  summarise(total_death = sum(cases), .groups = 'drop') %>%
  group_by(country) %>%
  mutate(week_movavg = zoo::rollmean(total_death, k = 7, fill = "extend")) %>%
  mutate(id = row_number()) %>% 
  filter(id %% 7 == 0) %>% # reduce the number of dots to plot to make ploting faster
  ungroup() %>% 
  {.}
```

```{r}
mov_avg_covid %>% 
  filter(id %% 7 == 0) %>% # reduce the number of dots to plot to make ploting faster
  ggplot(aes(date, week_movavg)) +
  geom_path() 
```

What happened?
How do we tell ggplot to draw lines between points of the same country?

```{r}
mov_avg_covid %>% 
  filter(id %% 7 == 0) %>% # reduce the number of dots to plot to make ploting faster
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path()
```

How can we tell which country is which?
Can we try colors?

```{r}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country, color = country)) +
  geom_path()
```

OK, this is cool, but perhaps too many colors.
How about we highlight only a few countries?

```{r}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, color = country)) +
  geom_path() +
  gghighlight(country %in% c('US', 'Korea, South'))
```

What about facets? (or small multiples)
```{r}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg)) +
  geom_path(size = 1.5) +
  facet_wrap(~ country)
```

*Tip*: Like piping into {.}, `+ NULL` allows you to comment out any line 
of the ggplot code (including the last one).

But I want to compare one country to the others...
```{r}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path() +
  facet_wrap(~ country) +
  gghighlight() +
  NULL
```

## PART 3: Fine tuning
### Exercises

1. Can we add `scale_y_log10()` to plot the y-axis in log scale? 
Do you think this is "better"?
Can we make the "labels" fancier by 
adding `labels = scales::comma_format()` inside `scale_y_log10()`?
```{r eval=FALSE}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path() +
  facet_wrap(~ country) +
  gghighlight() +
  ____________
```

2. Examine the `scale_x_date()` function to see if we can adjust the `x` ticks 
to "break" every 2 months? 
(Hint: run `?scale_x_date` on the console or a new chunk)

```{r eval=FALSE}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path() +
  facet_wrap(~ country) +
  gghighlight() +
  scale_x_date(breaks = ________)
```

3. Maybe add `ft_theme()`? (already sourced from `utils.R` earlier)
What about other gg_themes()?
```{r eval=FALSE}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path() +
  facet_wrap(~ country) +
  gghighlight() +
  scale_x_date(breaks = ________)
```

4. Do you think the label for the `x` axis is necessary? 
What about the `y` axis? What if we want to remove them?
```{r eval=FALSE}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path() +
  facet_wrap(~ country) +
  gghighlight() +
  labs(x = _________, y = _________) 
```

5. Can we make `#0A508C` the color of the highlighted country?
```{r eval=FALSE}
mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path(color = _________, size = 1.5) +
  gghighlight() +
  facet_wrap(~ country) +
  NULL
```

6. Putting 1-5 all together now, and let's save it to `figs/`.

```{r eval=FALSE}
ft_rep <- mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path(color = _________, size = 1.5) +
  gghighlight() +
  facet_wrap(~ country) +
  scale_y_log10(labels = _________) +
  scale_x_date(breaks = ________) +
  labs(x = _________, y = _________,
       title = _________,
       subtitle = _________) +
  ft_theme() +
  NULL

ft_rep
ggsave('figs/my_ft_rep.png', ft_rep, height = 10, width = 15)
```

Extra credits:
7. Can we use the argument `unhighlighted_params` in `gghighlight()` to make the unhighlighted countries thinner and lighter? (so that the highlighted country stands out  more)
8. Can we use the argument `sec.axis` in `scale_y_log10()` together with the function `sec_axis()` to add a secondary axis?
9. Can we try other conditions for filtering (from the beginning):
- Visualize only Asian countries?
- Visualize number of cases instead of death tolls?

```{r}

```

That's it!!!
These only few lines of code got you close to this chart:
https://twitter.com/jburnmurdoch/status/1250538655707430913,
except we plotted these with updated data!!!
Note slight difference between their `x` axis and ours (I did this intentionally so we could play with `scale_x_date`), 
extra end point at the end of each line (can you try adding this?), 
font (ggplot2 does not do great with outside fonts in my experience),
direct labeling, etc.
But you know what I mean :).


## PART 4: Let's see your charts!

Now that we're confident we could regenerate any chart, let's play around with this one a bit more.
You can change the theme (check out `library(ggthemes)`), color of the highlighted countries, 
or even put them back into just one "big single" rather than "small multiples",
add some direct labeling with `geom_text()` or `library(ggrepel)`,
etc.
(Check out https://ig.ft.com/coronavirus-chart/ for inspiration!)
Please share your final chart(s) here:
https://drive.google.com/drive/folders/1N7VNDqBkakzX07XZaiO5xUKtqdqLbX8z?usp=sharing


References:
https://peymankor.netlify.app/post/ft/
https://www.ft.com/content/a2901ce8-5eb7-4633-b89c-cbdf5b386938





























































































































```{r}
p <- mov_avg_covid %>% 
  ggplot(aes(date, week_movavg, group = country)) +
  geom_path(color = '#0A508C', lineend = "round", size = 1.5) +
  gghighlight(unhighlighted_params = list(size = 0.5, colour = alpha("grey", 0.4))) +
  facet_wrap(~ country) +
  scale_y_log10(labels = scales::comma_format(accuracy = 1),
                sec.axis = sec_axis(~ ., labels = scales::comma_format(accuracy = 1))) +
  scale_x_date(breaks = '2 months', date_labels = '%b', 
               expand = expansion(add = 0, mult = 0)) +
  labs(x = NULL, y = NULL,
       title = "Daily death tolls",
       subtitle = "Daily death with coronavirus (7-day rolling average), by number of the days since 3 daily deaths first recorded") +
  ft_theme() +
  NULL

ggsave('ft_rep.png', p, height = 10, width = 15)
```
